---
title: "Catching up with R"
author: "Econ 695"
date: "Spring 2024"
output: pdf_document
---

## Load the data and the packages

To begin with, load the data (and the packages necessary for your work).

```{r, message = FALSE}
# load packages
library(tidyverse)
library(stargazer)
# load data
resume <- read.csv("resume.csv")
```


## Check the data

You can use view() function or simply double-click the dataframe on the right panel.

```{r}
# How many observations are female?
table(resume$sex)
# How many observatles are female and white?
table(resume$sex,resume$race)
```


## Create variables

```{r}
# generate a dummy for female
resume$female <- 0
resume$female[resume$sex=="female"] <- 1
resume$female <- as.factor(resume$female)
```


## Comparing means

```{r}
# Are callback rates different by race?
tapply(resume$call, resume$race, mean)
summarize(group_by(resume,race),mean(call))
# t-test
t.test(resume$call[resume$race=="white"],resume$call[resume$race=="black"])
# t-stat?
t.test(resume$call[resume$race=="white"],resume$call[resume$race=="black"])$statistic
```


## Running regression

```{r}
# Are callback rates differt by race? (reprise)
fit <- lm(call~race, data = resume)
summary(fit)
# Is the effect of race on callback rates different by gender?
fit2 <- lm(call~race*sex, data = resume)
summary(fit2)
```


## Stargzer

Using the stargazer package, you can present multiple regression results in a table, in a compact manner.

```{r, message = FALSE, warning = FALSE}
stargazer(fit, fit2, type="text")
stargazer(fit, fit2, type= "text",
          title = "Effects of Race on Callback rates",
          covariate.labels = c("White", "Male", "White x Male"),
          dep.var.caption = "Dependent variable: Received a callback or not",
          dep.var.labels.include = FALSE,
          model.names = NULL,
          omit.stat = c("ser", "f"),
          omit = c("Constant"),
          notes = "More notes.", notes.align = "r",
          header = FALSE)
```